---
title: "MDS DGE analysis"
author: "Monica Ransom"
date: "8/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-processing
1.    Adapters were trimmed using cutadapt v1.16
2.    Gene expression was quantified using salmon  v1.3.0
3.    TPMs were obtained for the genes using tximport 1.20.0

```{r import, message=FALSE}
library(dplyr)
library(ggplot2)
library(DESeq2)
library(tximport)
library(readr)
library(tximportData)
library(readxl)
library(knitr)
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
library(viridis)
library(ggrepel)
library(EnhancedVolcano)
library(fgsea)
library(limma)
```
## R Markdown

## Define Color Scheme and plot them
```{r, echo=FALSE, include=FALSE}
plasma <- viridis::plasma(n = 10)
viridis <- viridis::viridis(n = 10)
cividis <- viridis::cividis(n=10)
magma <- viridis::magma(n=10)
```
<br>
___Plasma 1:10___
```{r, echo=FALSE, fig.width=5, fig.height=1}
qplot(x=1:length(plasma), y = 1, fill=factor(1:length(plasma)), geom="tile") +
  scale_fill_manual(values = plasma) +
  theme_void()+
  theme(legend.position="none") 
```
<br>
___Viridis 1:10___
```{r, echo=FALSE, fig.width=5, fig.height=1}
qplot(x=1:length(viridis), y = 1, fill=factor(1:length(plasma)), geom="tile") +
  scale_fill_manual(values = viridis) +
  theme_void()+
  theme(legend.position="none") 
```
<br>
___Cividis 1:10___
```{r, echo=FALSE, fig.width=5, fig.height=1}
qplot(x=1:length(cividis), y = 1, fill=factor(1:length(plasma)), geom="tile") +
  scale_fill_manual(values = cividis) +
  theme_void()+
  theme(legend.position="none") 
```
<br>
___Magma 1:10___
```{r, echo=FALSE, fig.width=5, fig.height=1}
qplot(x=1:length(magma), y = 1, fill=factor(1:length(plasma)), geom="tile") +
  scale_fill_manual(values = magma) +
  theme_void()+
  theme(legend.position="none") 
```

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. 

##This section of code will generate a summary table of the samples sequenced and their sequencing and alignment metrics
```{r summary, echo=FALSE}
MDS_metric <- read_excel("~/Desktop/Jordan files/Results/MDS stats.xlsx")

starting_metrics<-filter(MDS_metric,grepl('HTB314|MDS268|MDS280', patient))
kable(starting_metrics, caption="Summary of Data Metrics")
```

##This section of code will import and format the data for DeSeq2

```{r data, message=FALSE}
#to generate a vector of names and file locations
files<-file.path("~/Desktop/Jordan files/Counts/salmon/salmon/", list.files("~/Desktop/Jordan files/Counts/salmon/salmon/"), "quant.sf")
names(files)<-list.files("~/Desktop/Jordan files/Counts/salmon/salmon")

#to call in a gene_map this was derived by pulling it out of the fasta file with a grep command for ENST and ENSG and pasting them together
gene_map=read_csv("~/Desktop/Jordan files/Counts/salmon/gene_map.csv", col_names=c('enstid', 'ensgid'))

# import transcript level counts
txi.salmon.t<-tximport(files, type="salmon", txOut=TRUE)
txi.salmon.g<-tximport(files=files, type="salmon", tx2gene= gene_map, ignoreTxVersion = TRUE, countsFromAbundance = 'lengthScaledTPM' )
### this code works but if I remove the ignoreTxVersion I get an error, this may have to do with how I am generating my tx2gene file

# Extract counts only
counts <- txi.salmon.g$counts %>%
  as.data.frame()
#Extract TPM
tpms <- data.frame(txi.salmon.g$abundance)

##for clients the counts and tpm files should be written out

```

##This chunk of code sets an expression threshold for TMP where all samples have at least 5 counts
```{r expressed genes}

expressed_genes<-tpms %>% filter_all(all_vars(. > 5))
```

### This code is to bring in the .sf files that have the 45s rRNA included.  This will be used as a metric to include in our PCA plot.
```{r}
files.rna<-file.path("~/Desktop/Jordan files/counts.rRNA/rRNA/", list.files("~/Desktop/Jordan files/counts.rRNA/rRNA"), "quant.sf")
names(files.rna)<-list.files("~/Desktop/Jordan files/counts.rRNA/rRNA")

mart <- biomaRt::useMart("ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host='www.ensembl.org')
t2g_hs <- biomaRt::getBM(attributes = c('ensembl_transcript_id', 'ensembl_gene_id', 'external_gene_name', 'refseq_mrna'), mart = mart)
tx2gene <- t2g_hs[,c(1,2)]
tx2gene[nrow(tx2gene)+1,]=c('rRNA_45S_NR145819','rRNA_45S_NR145819')

txi.salmon.rRNA.g<-tximport(files=files.rna, type="salmon", tx2gene= tx2gene, ignoreTxVersion = TRUE, countsFromAbundance = 'lengthScaledTPM' )
metadata<-read_table2("~/Desktop/Jordan files/Counts/sample.txt")
sums<-colSums(txi.salmon.rRNA.g$counts)
sums<-as.data.frame(sums)
counts.rRNA<-as.data.frame(txi.salmon.rRNA.g$counts)
rna.counts<-as.data.frame(t(slice_tail(counts.rRNA)))
rna.counts <-merge(rna.counts,sums,by='row.names',all=TRUE)
rna.counts<-mutate(rna.counts, "percentrRNA"= (rRNA_45S_NR145819 /sums)*100 )
names(rna.counts)[1] <- 'SampleName'
rna.counts<-select(rna.counts, c("SampleName", "percentrRNA"))
metadata<-dplyr::full_join(metadata,rna.counts, by="SampleName")
metadata$percentrRNA <- round(metadata$percentrRNA ,digit=2)

```

##This chunk of code generates a heatmap using the spearman method as well as a correlation heatmap
```{r All samples Heatmap, echo=FALSE, fig.width=7.5, fig.height=7, fig.align='center'}
## what we see here shows that patients cluster more closely than the cell types and this will need to be accounted for in the DGE
pheatmap(expressed_genes, scale="row")
cols.cor<-cor(expressed_genes, method="spearman")

pheatmap(cols.cor, method="spearman")
```
## PCA plot



```{r pca plot, message=FALSE}
exp.pca <- prcomp(t(log2(expressed_genes)))
exp.pca.summary <- summary(exp.pca)$importance
pc1var = round(exp.pca.summary[3,1] * 100, 1)
pc2var = round(exp.pca.summary[3,2] * 100 - pc1var, 1)
exp.pca.pc <- data.frame(exp.pca$x, sample = colnames(expressed_genes))

```

### PC1 vs PC2
```{r, All samples PCA 1 vs 2, echo = FALSE, warning=FALSE,  fig.width= 6, fig.height=4.5, fig.align='center'}

colors <- c(viridis[6], cividis[9], magma[6])
ggplot(exp.pca.pc, aes(x = PC1, y = PC2, color=metadata$cellType, shape=metadata$patient)) +
  geom_point(size = 5)+
  scale_fill_manual(values = alpha(colors, 1) ) + 
  xlab(paste('PC1,', pc1var, '% explained var.')) +
  ylab(paste('PC2,', pc2var, '% explained var.')) +
  guides(fill=guide_legend(override.aes = list(color=colors))) +
  geom_text_repel(aes(label=metadata$SampleName), size=3, max.overlaps = Inf)+
  ggtitle("Expressed Genes in All Samples > 5 TPM") 
```
## 3. Run Differential Expression testing using DESeq2 and Calculate Gene Set Enrichment
## Compare 123pos vs 123neg, 123neg vs bulk, and 123pos vs bulk 
#### sig = padj <0.01 and abs(l2fc) >0.5
#### 

```{r DGE}
coldata<-dplyr::select(metadata, -FileName) 
coldata<-column_to_rownames(coldata, 'SampleName')
## need to confirm that all names are in the same order
all(rownames(coldata) %in% colnames(txi.salmon.g$counts))
all(rownames(coldata) == colnames(txi.salmon.g$counts))
coldata<-coldata[colnames(txi.salmon.g$counts),]
all(rownames(coldata) == colnames(txi.salmon.g$counts))

dds <- DESeqDataSetFromTximport(txi.salmon.g,
                              colData = coldata,
                              design = ~ patient + cellType)

dds<-estimateSizeFactors(dds)
sf<-as.data.frame(dds$sizeFactor)
sf<-rownames_to_column(sf, var="SampleName")
metadata<-inner_join(metadata, sf, by="SampleName")
names(metadata)[6] <- "sizeFactor"
metadata$sizeFactor <- round(metadata$sizeFactor ,digit=2)
### unfiltered data
dds.unfiltered <- DESeq(dds)
res.unfiltered <- results( dds.unfiltered)

##filtered data changed this from dds[rowMins(counts(dds))>5,] to a less stringent filtering
keep<-rowSums(counts(dds))>=10
#dds.filtered<-dds[rowMins(counts(dds))>5,]
dds.filtered<-dds[keep,]
dds.filtered<-DESeq(dds.filtered)
res.filtered<-results(dds.filtered)

###outcome the first filtering cut from 60000 genese to 14000 genes with the new filter 35000genes are left which is about 1/2
### need to subset to do pairwise comparison unclear if this keeps the patient design aspect


res_123posvs123neg_unfiltered<-results( dds.unfiltered, contrast=c("cellType", "123pos", "123neg"))
res_123posvsBulk_unfiltered<-results( dds.unfiltered, contrast=c("cellType", "123pos", "bulk"))
res_123negvsBulk_unfiltered<-results( dds.unfiltered, contrast=c("cellType", "123neg", "bulk"))

res_123posvs123neg_filtered<-results(dds.filtered, contrast=c("cellType", "123pos", "123neg"))
res_123posvsBulk_filtered<-results(dds.filtered, contrast=c("cellType", "123pos", "bulk"))
res_123negvsBulk_filtered<-results(dds.filtered, contrast=c("cellType", "123neg", "bulk"))

## look at the numbers of genes meeting threshold the log fold change call is not changing things
sum( res_123posvs123neg_unfiltered$pvalue < 0.01 & abs(res_123posvs123neg_unfiltered$log2FoldChange) >= 0.5, na.rm=TRUE )
sum( res_123posvsBulk_unfiltered$pvalue < 0.01 & abs(res_123posvsBulk_unfiltered$log2FoldChange) >= 0.5, na.rm=TRUE )
sum(res_123negvsBulk_unfiltered$pvalue < 0.01 & abs(res_123negvsBulk_unfiltered$log2FoldChange) >= 0.5, na.rm=TRUE )

sum( res_123posvs123neg_filtered$pvalue < 0.01 & abs(res_123posvs123neg_filtered$log2FoldChange) >= 0.5, na.rm=TRUE )
sum( res_123posvsBulk_filtered$pvalue < 0.01 & abs(res_123posvsBulk_filtered$log2FoldChange) >= 0.5, na.rm=TRUE )
sum(res_123negvsBulk_filtered$pvalue < 0.01 & abs(res_123negvsBulk_filtered$log2FoldChange) >= 0.5, na.rm=TRUE )

sum(res_123posvs123neg_filtered$padj < 0.1, na.rm=TRUE)
sum(res_123posvs123neg_unfiltered$padj < 0.1, na.rm=TRUE)
```


### Volcano Plot
```{r, echo=FALSE, fig.width= 10, fig.height=6}

colors <- c("black", viridis[5])
df<-data.frame(res_123posvs123neg_filtered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123pos vs 123neg filtered")

colors <- c("black", viridis[5])
df<-data.frame(res_123posvs123neg_unfiltered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123pos vs 123neg unfiltered")

colors <- c("black", viridis[5])
df<-data.frame(res_123posvsBulk_unfiltered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123pos vs Bulk unfiltered")

colors <- c("black", viridis[5])
df<-data.frame(res_123posvsBulk_filtered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123pos vs bulk filtered")
 
colors <- c("black", viridis[5])
df<-data.frame(res_123negvsBulk_unfiltered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123neg vs Bulk unfiltered")

colors <- c("black", viridis[5])
df<-data.frame(res_123negvsBulk_filtered)
df<-mutate(df, DiffExp=pvalue<=.05 & abs(log2FoldChange) >=.5)
ggplot(df, aes(x=log2FoldChange, y=-log10(padj), color=DiffExp)) + 
  geom_point() +
  theme_light() +
  scale_colour_manual(values = colors) +
  ggtitle("123neg vs bulk filtered")

```

###MA plots
```{r, echo=FALSE, fig.width= 6, fig.height=4.5}

DESeq2::plotMA(res_123posvs123neg_unfiltered, main="123pos vs 123neg unfiltered", 
       ylim=c(-5,5))
DESeq2::plotMA(res_123posvsBulk_unfiltered,  main="123pos vs Bulk unfiltered", 
       ylim=c(-5,5))
DESeq2::plotMA(res_123negvsBulk_unfiltered,  main="123neg vs Bulk unfiltered", 
       ylim=c(-5,5))

DESeq2::plotMA(res_123posvs123neg_filtered,  main="123pos vs 123neg filtered", 
       ylim=c(-5,5))
DESeq2::plotMA(res_123posvsBulk_filtered,  main="123pos vs Bulk filtered", 
       ylim=c(-5,5))
DESeq2::plotMA(res_123negvsBulk_filtered,  main="123neg vs Bulk filtered", 
       ylim=c(-5,5))
```



```{r pca plot 2, message=FALSE}
pca_data<-vst(dds, blind=T)
ntop=500
rv <- rowVars(assay(pca_data))
select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]
pca <- prcomp(t(assay(pca_data)[select,]))
percentVar2 <- data.frame(percentVar=pca$sdev^2/sum(pca$sdev^2))%>%
  mutate(pc=1:n())%>%dplyr::select(pc, percentVar)
pca_df<-pca$x%>%data.frame()%>%mutate(type=metadata$cellType,percent=metadata$percentrRNA)

alt_col_values=c("#88CCEE", "#CC6677", "#DDCC77")

ggplot(pca_df, aes(PC1, PC2, color=type, shape=metadata$patient))+geom_point(size=3)+
  xlab(paste0("PC1: ",round(percentVar2$percentVar[1] * 100), "% variance")) + 
  ylab(paste0("PC2: ", round(percentVar2$percentVar[2] * 100), "% variance"))+
  scale_color_manual(values=alt_col_values)+labs(color="")+theme_bw()+
  ggtitle("PCA of analyzed data")
```




```{r limma}
pca_data_limma<-limma::removeBatchEffect(assay(pca_data), pca_data$patient )

#rv <- rowVars(assay(pca_data_limma))
#select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]
pca_limma <- prcomp(t(pca_data_limma))
percentVar3 <- data.frame(percentVar=pca_limma$sdev^2/sum(pca_limma$sdev^2))%>%
  mutate(pc=1:n())%>%dplyr::select(pc, percentVar)
pca_df_limma<-pca_limma$x%>%data.frame()%>%mutate(type=metadata$cellType)

ggplot(pca_df_limma, aes(PC1, PC2, color=type, shape=metadata$patient))+geom_point(size=3)+
  xlab(paste0("PC1: ",round(percentVar3$percentVar[1] * 100), "% variance")) + 
  ylab(paste0("PC2: ", round(percentVar3$percentVar[2] * 100), "% variance"))+
  scale_color_manual(values=alt_col_values)+labs(color="")+theme_bw()+
  ggtitle("PCA of analyzed data with limma")
```

### PC1 vs PC2




```{r pca plot 3, message=FALSE}
pca_data2<-vst(dds.filtered, blind=T)
ntop=500
rv <- rowVars(assay(pca_data2))
select <- order(rv, decreasing=TRUE)[seq_len(min(ntop, length(rv)))]
pca2 <- prcomp(t(assay(pca_data2)[select,]))
percentVar4 <- data.frame(percentVar=pca2$sdev^2/sum(pca2$sdev^2))%>%
  mutate(pc=1:n())%>%dplyr::select(pc, percentVar)
pca_df2<-pca2$x%>%data.frame()%>%mutate(type=metadata$cellType, percent=metadata$percentrRNA)

alt_col_values=c("#88CCEE", "#CC6677", "#DDCC77")

ggplot(pca_df2, aes(PC1, PC2, color=type, shape=metadata$patient))+geom_point(size=3)+
  xlab(paste0("PC1: ",round(percentVar4$percentVar[1] * 100), "% variance")) + 
  ylab(paste0("PC2: ", round(percentVar4$percentVar[2] * 100), "% variance"))+
  scale_color_manual(values=alt_col_values)+labs(color="")+theme_bw()+
  ggtitle("PCA of analyzed data using filtered data")
```


```{r, pca 4 plot, echo = FALSE, fig.width= 6, fig.height=4.5}

ggplot(pca_df2, aes(PC1, PC2, color=type, shape=metadata$patient, label=metadata$sizeFactor))+geom_point(size=3)+
  geom_text(aes(label=metadata$sizeFactor),hjust=0, vjust=0)+
  xlab(paste0("PC1: ",round(percentVar3$percentVar[1] * 100), "% variance")) + 
  ylab(paste0("PC2: ", round(percentVar3$percentVar[2] * 100), "% variance"))+
  scale_color_manual(values=alt_col_values)+labs(color="")+theme_bw()+
  ggtitle("PCA of analyzed data labelled with size factor")
```

```{r, pca 5 plot, echo = FALSE, fig.width= 6, fig.height=4.5}

ggplot(pca_df2, aes(PC1, PC2, color=type, shape=metadata$patient, label=metadata$percentrRNA))+geom_point(size=3)+
  geom_text(aes(label=metadata$percentrRNA),hjust=0, vjust=0)+
  xlab(paste0("PC1: ",round(percentVar3$percentVar[1] * 100), "% variance")) + 
  ylab(paste0("PC2: ", round(percentVar3$percentVar[2] * 100), "% variance"))+
  scale_color_manual(values=alt_col_values)+labs(color="")+theme_bw()+
  ggtitle("PCA of analyzed data labelled with rRNA percentage")
```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
##generates a plot of FBX015 a gene in paper up in 123+
d<-plotCounts(dds, gene="ENSG00000141665",intgroup=c("cellType","patient"), returnData = TRUE)

ggplot(d, aes(x=cellType, y=count, color=patient))+geom_point()
```

```{r}
##generates a plot of EEFSEC a gene in paper up in 123+
d<-plotCounts(dds, gene="ENSG00000132394",intgroup=c("cellType","patient"), returnData = TRUE)

ggplot(d, aes(x=cellType, y=count, color=patient))+geom_point()
```

```{r}
##generates a plot of TGM4 a gene in paper down in 123+
d<-plotCounts(dds, gene="ENSG00000163810",intgroup=c("cellType","patient"), returnData = TRUE)

ggplot(d, aes(x=cellType, y=count, color=patient))+geom_point()
```


```{r}
##generates a plot of CRLF2 a gene in paper down in 123+
d<-plotCounts(dds, gene="ENSG00000163810",intgroup=c("cellType","patient"), returnData = TRUE)

ggplot(d, aes(x=cellType, y=count, color=patient))+geom_point()
```

###GSEA analysis
```{r GSEA}
##generate gene names to go with ensembl gene id as the pathways will use gene name
mart <- biomaRt::useMart("ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host='www.ensembl.org')
t2g_hs <- biomaRt::getBM(attributes = c('ensembl_transcript_id', 'ensembl_gene_id', 'external_gene_name', 'refseq_mrna'), mart = mart)
ens2gene <- t2g_hs[,c(2,3)]
colnames(ens2gene)[2] <- 'Gene'
ens2gene <- unique(ens2gene)

##loading hallmark pathways and KEGG pathways

pathways.hallmark <- gmtPathways("~/Desktop/Jordan files/h.all.v7.4.symbols.gmt")
pathways.kegg<-gmtPathways("~/Desktop/Jordan files/c2.cp.kegg.v7.4.symbols.gmt")

##generating tidy data and adding it to the results file
##123pos vs 123neg analysis
res_posneg_gsea<-results(dds.filtered, contrast=c("cellType", "123pos", "123neg"),tidy=TRUE)
colnames(res_posneg_gsea)[1]<-"ensembl_gene_id"
res_posneg_gsea <- inner_join(res_posneg_gsea,ens2gene,by="ensembl_gene_id")
res_posneg_gsea2<-res_posneg_gsea%>%
  dplyr::select(Gene, stat) %>%
  na.omit() %>%
  distinct() %>%
  group_by(Gene) %>%
  summarise(stat=mean(stat))
rank_posneg<-deframe(res_posneg_gsea2)
fgseaRes_posneg <- fgsea(pathways=pathways.hallmark, stats=rank_posneg)
fgseaRes_posneg_kegg<-fgsea(pathways=pathways.kegg, stats=rank_posneg)

fgseaRes_posneg_tidy <-fgseaRes_posneg%>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaRes_posneg_kegg_tidy <-fgseaRes_posneg_kegg%>%
  as_tibble() %>%
  arrange(desc(NES))

##123 pos vs bulk analysis

res_posbulk_gsea<-results(dds.filtered, contrast=c("cellType", "123pos", "bulk"),tidy=TRUE)
colnames(res_posbulk_gsea)[1]<-"ensembl_gene_id"
res_posbulk_gsea <- inner_join(res_posbulk_gsea,ens2gene,by="ensembl_gene_id")

res_posbulk_gsea2<-res_posbulk_gsea%>%
  dplyr::select(Gene, stat) %>%
  na.omit() %>%
  distinct() %>%
  group_by(Gene) %>%
  summarise(stat=mean(stat))
rank_posbulk<-deframe(res_posbulk_gsea2)
fgseaRes_posbulk<- fgsea(pathways=pathways.hallmark, stats=rank_posbulk)
fgseaRes_posbulk_kegg<-fgsea(pathways=pathways.kegg, stats=rank_posbulk)

fgseaRes_posbulk_tidy <-fgseaRes_posbulk%>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaRes_posbulk_kegg_tidy <-fgseaRes_posbulk_kegg%>%
  as_tibble() %>%
  arrange(desc(NES))

##123 neg vs bulk analysis
res_negbulk_gsea<-results(dds.filtered, contrast=c("cellType", "123neg", "bulk"),tidy=TRUE)
colnames(res_negbulk_gsea)[1]<-"ensembl_gene_id"
res_negbulk_gsea <- inner_join(res_negbulk_gsea,ens2gene,by="ensembl_gene_id")


res_negbulk_gsea2<-res_negbulk_gsea%>%
  dplyr::select(Gene, stat) %>%
  na.omit() %>%
  distinct() %>%
  group_by(Gene) %>%
  summarise(stat=mean(stat))
rank_negbulk<-deframe(res_negbulk_gsea2)
fgseaRes_negbulk<- fgsea(pathways=pathways.hallmark, stats=rank_negbulk)
fgseaRes_negbulk_kegg<-fgsea(pathways=pathways.kegg, stats=rank_negbulk)

fgseaRes_negbulk_tidy <-fgseaRes_negbulk%>%
  as_tibble() %>%
  arrange(desc(NES))

fgseaRes_negbulk_kegg_tidy <-fgseaRes_negbulk_kegg%>%
  as_tibble() %>%
  arrange(desc(NES))
```

```{r}
fgseaRes_posneg_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: Hallmark genes with 123pos vs 123neg.')
```

```{r}
fgseaRes_posneg_kegg_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: KEGG pathways with 123pos vs 123neg.')
```
```{r}
fgseaRes_posbulk_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: Hallmark genes with 123pos vs bulk.')
```

```{r}
fgseaRes_posbulk_kegg_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: KEGG pathways with 123pos vs bulk.')
```

```{r}
fgseaRes_negbulk_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: Hallmark genes with 123neg vs bulk.')
```

```{r}
fgseaRes_posneg_kegg_tidy%>%
  dplyr::select(-leadingEdge, -ES) %>%
  arrange(padj) %>%
  DT::datatable(caption = 'Table 1: KEGG pathways with 123neg vs bulk.')
```

```{r, fig.width= 6, fig.height=8}
ggplot(fgseaRes_posneg_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Hallmark pathways NES from GSEA 123pos vs 123neg") + 
  theme_minimal()
```
```{r, fig.width= 8, fig.height=16}
ggplot(fgseaRes_posneg_kegg_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Kegg pathways NES from GSEA 123pos vs 123neg") + 
  theme_minimal()
```

```{r, fig.width= 6, fig.height=8}
ggplot(fgseaRes_posbulk_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Hallmark pathways NES from GSEA 123pos vs bulk") + 
  theme_minimal()
```

```{r, fig.width= 8, fig.height=16}
ggplot(fgseaRes_posbulk_kegg_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Kegg pathways NES from GSEA 123pos vs bulk") + 
  theme_minimal()
```

```{r, fig.width= 6, fig.height=8}
ggplot(fgseaRes_negbulk_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Hallmark pathways NES from GSEA 123neg vs bulk") + 
  theme_minimal()
```

```{r, fig.width= 8, fig.height=16}
ggplot(fgseaRes_negbulk_kegg_tidy, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=padj<0.1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Kegg pathways NES from GSEA 123neg vs bulk ") + 
  theme_minimal()
```